---
title: "Sexual Offenses and Alcohol Abuse at Pennsylvania State University Campuses"
author: "Sarah Kircher, Leigh Stern, and Nathanael Kovscek"
date: "`r Sys.Date()`"
output: html_notebook
---

# Introduction
Insert introduction here.

```{r setUp}
# Load needed packages
library(ggplot2)
library(readxl)

```

# Population Analysis
Insert population plots and analysis here.

# University Park Cases

Explain Penn State PULSE survey here.

```{r alcData}
# Load dataset 1 from Excel - PSU student alcohol consumptions
alcohol_consumption <- read_excel("PSU_students_report_consuming_alc.xlsx")

# Load dataset 2 from Excel - PSU student high risk drinking by sex
hr_drinking_sex <- read_excel("PSU_high_risk_drinking_sex.xlsx")

# Load dataset 3 from Excel - PSU student drinking consequences
drinking_consequences <- read_excel("PSU_drinking_consequences.xlsx")
```

## Question
```{r alcoholConsumption}
# Create ggplot2 for PSU student alcohol consumption over time
# One line/trend for respondents who do drink
# One line/trend for respondents who do not drink
alc_consump_colors <- c("Drinker" = "red", "Non-Drinker" = "gray")

ggplot(
  data = alcohol_consumption, 
  aes(x = year)
) + 
  geom_point(aes(y = drink_num, color = "Drinker"), size = 2.25) +
  geom_point(aes(y = not_drink_num, color = "Non-Drinker"), size = 2.25) +
  
  geom_line(aes(y = drink_num, color = "Drinker"), size = 1) + 
  geom_line(aes(y = not_drink_num, color = "Non-Drinker"), size = 1) +
  labs(x = "Year", y = "Respondents", color = "Drinker Status") +
  ggtitle("Penn State Student Alcohol Consumption Over Time") +
  scale_color_manual(values = alc_consump_colors) +
  theme_minimal()

```
Analyze plot here.

## Second Question
```{r highriskDrinking}
# Creates grouped bar chart for PSU high risk drinking by sex over time
# One color bar (blue) for male respondents
# One color bar (pink) for female respondents
ggplot(
  data = hr_drinking_sex, 
  aes(x = year,
      y = frequency,
      fill = sex)) +
  geom_bar(stat = "identity", position = 'dodge') +
  ggtitle("Penn State Student High Risk Drinking By Sex Over Time") +
  labs(x = "Year", y = "Respondents", fill = "Sex") +
  scale_fill_manual(values = c("pink", "blue")) +
  theme_minimal()

```

Analyze plot here.

## Third Question
```{r drinkingConsequences}
# Create ggplot2 for drinking consequences over time
# One line for "Had sex when you didn’t really want to"
# One line for "Been pressured or forced to have sex with someone and were too drunk to prevent it"
# One line for "Pressured or forced someone to have sex with you after you had been drinking"
consequence_colors <- c("did not want to" = "dark green", 
                                 "were pressured when too drunk" = "purple",
                                 "were pressured after drinking" = "orange")
ggplot(
  data = drinking_consequences, 
  aes(x = year)
) + 
  geom_point(aes(y = sex_did_not_want, color = "did not want to"), size = 2.25) +
  geom_point(aes(y = pressured_forced_drunk, color = "were pressured when too drunk"), size = 2.25) +
  geom_point(aes(y = pressured_forced_after_drinking, color = "were pressured after drinking"), size = 2.25) +
  
  geom_line(aes(y = sex_did_not_want, color = "did not want to"), size = 1) + 
  geom_line(aes(y = pressured_forced_drunk, color = "were pressured when too drunk"), size = 1) +
  geom_line(aes(y = pressured_forced_after_drinking, color = "were pressured after drinking"), size = 1) +
  labs(x = "Year", y = "Percentage of Respondents", color = "Respondent had sex when they...") +
  ggtitle("Penn State Student Drinking Consequence Responses Over Time") +
  scale_color_manual(values = consequence_colors) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = c(2010, 2012, 2014, 2016, 2018, 2020)) +
  theme_minimal()

```

Analyze plot here.

# Branch Campus Cases
Insert table and branch campus information here.

# Synthesis
Insert synthesis here.

# Conclusion
Insert conclusion here.



